Incentivizing Exploration with Unbiased History

### 2018-10-02 2level.tex
Update 2-level proof

%!TEX root = main.tex
\section{Two-level Recommendation Policy}
Run $T_2$ \ALGG in parallel and then show the later agents the history of first $T_2$ rounds.

\begin{lemma}
For each arm $i$, any $t > 0$ and $\varepsilon \leq 1/2$, let $p$ be the probability that there exists $t \leq \tau \leq T$ such that the empirical mean of the first $\tau$ pulls of arm $i$ is $\varepsilon$ away from $\mu_i$. Then we have 
\[
p \leq \frac{2}{\varepsilon^2} \exp(-2\varepsilon^2 t).
\]
\end{lemma}

\begin{proof}
For each $\tau \in \{t,...,T\}$, let $p_{\tau}$ be the probability that the empirical mean of the first $\tau$ pulls of arm $i$ is $\varepsilon$ away from $\mu_i$. By Chernoff bound, we have
\[
p_{\tau} \leq 2 \exp(-2\varepsilon^2 \tau).
\]

By union bound,
\[
p \leq \sum_{\tau = t}^T p_{\tau} \leq 2 \exp(-2\varepsilon^2 t)  \cdot \frac{1}{1-\exp(-2\varepsilon^2)}.
\]

We know that for any $ x \in [0,1]$, $e^{-x} \leq 1-x/2$. There we have
\[
p \leq  2 \exp(-2\varepsilon^2 t)  \cdot \frac{1}{1-(1-\varepsilon^2)} =  \frac{2}{\varepsilon^2} \exp(-2\varepsilon^2 t).
\]
\end{proof}

\begin{theorem}
\2LEVEL gets expected regret $O(T^{2/3} \log(T)^{1/3})$. If $\delta$, the expected regret is $O(\log(T) \cdot \delta^{-2})$. 
\end{theorem}

\begin{proof}
We will analyze the case when agents after \ALGG only uses empirical mean.
\end{proof}

>>>>

%!TEX root = main.tex
\section{Warm-up: Two-level Recommendation Policy}
In this section, we are going to show a two-level recommendation policy that gets sublinear regret.
In this first level, we run $T_1$ \ALGG ``in parallel''. More specifically, we divide the first $T_1 \cdot \GdT$ agents into $T_1$ groups and agents only observe the history of all previous agents in the same group.
In the second level, we show the rest $T-T_1\cdot \GdT$ agents the history of first $T_1 \cdot \GdT$ rounds. We will pick $T_1$ in Theorem \ref{thm:2level}. 
See also Figure \ref{fig:2level} for a graphical view of the information flow.

\begin{figure}[H]
\centering
\begin{tikzpicture}  
 \filldraw[fill=blue!20!white]
 (0,2)--(10,2)--(10,3)--(0,3)--cycle;
  \filldraw[fill=red!20!white]
  (0,0)--(1,0)--(1,1)--(0,1)--cycle;
  \draw[dashed] (0.5,1)--(5,2);
  \filldraw[fill=red!20!white]
  (1,0)--(2,0)--(2,1)--(1,1)--cycle;
  \draw[dashed] (1.5,1)--(5,2);
  \filldraw[fill=red!20!white]
  (2,0)--(3,0)--(3,1)--(2,1)--cycle;
  \draw[dashed] (2.5,1)--(5,2);
  \filldraw[fill=red!20!white]
  (3,0)--(4,0)--(4,1)--(3,1)--cycle;
  \draw[dashed] (3.5,1)--(5,2);
  \filldraw[fill=red!20!white]
  (9,0)--(10,0)--(10,1)--(9,1)--cycle;
  \draw[dashed] (9.5,1)--(5,2);
  \node at(5,0.5){$\cdots$};
  \node at(6,0.5){$\cdots$};
  \node at(7,0.5){$\cdots$};
  \node at(8,0.5){$\cdots$};
  \node at(5,2.5){$T -T_1 \cdot \GdT$};
  \node at(0.5,0.5){$\GdT$};
  \node at(1.5,0.5){$\GdT$};
  \node at(2.5,0.5){$\GdT$};
  \node at(3.5,0.5){$\GdT$};
  \node at(9.5,0.5){$\GdT$};
  \node at(-2,0.5){\textbf{First Level}};
  \node at(-2,2.5){\textbf{Second Level}};
\end{tikzpicture}
\caption{2-level Recommendation Policy.}
\label{fig:2level}
\end{figure}

\begin{theorem}
\label{thm:2level}
With proper setting of $T_1$, our two-level recommendation policy gets expected regret $O(T^{2/3} \log(T)^{1/3})$. If we know the gap between the means of the best arm and second best arm is larger than $\Delta$, we can make the expected regret is $O(\log(T) \cdot \Delta^{-2})$. 
\end{theorem}

\begin{proof}
We will set $T_1$ later in the proof. For now, we just need to make $T_1 \geq \frac{4\GdT^2}{\GdP^2}\log(T)$.

The first $T_1 \cdot \GdT$ rounds will get total regret at most $T_1 \cdot \GdT$.  We focus on the regret of the rest $T - T_1 \cdot \GdT$ rounds. We first consider the following two clean events and we show that they together imply upper bounds on $|\hat{\mu}^t_a - \mu_a|$'s for $t > T_1 \cdot \GdT$.

\begin{itemize}
\item For each arm $a$, define $q_a$ to be the expected number of arm $a$ pulls in one run of \ALGG used in the first level. By Lemma \ref{lem:greedy}, we know $\GdP \leq q_a \leq \GdT$ Define $W_1^a$ to be the event that the number of arm $a$ pulls in the first level is at least $q_a T_1- \GdT \sqrt{T_1\log(T)} \geq q_a T_1 / 2$. By Chernoff bound,
\[
\Pr[W_1^a] \geq 1-\exp(-2\log(T)) \geq 1-1/T^2.
\]
Define $W_1$ to be the intersection of all these events (i.e. $W_1 = \bigcap_{a}W_1^a$). By union bound, we have
\[
\Pr[W_1] \geq 1- \frac{K}{T^2} \geq 1 - \frac{1}{T}.
\]
\item For each arm $a$, imagine there is a tape of $T$ arm $a$ pulls sampled before the recommendation policy starts and these samples are revealed one by one whenever agent pull arm $a$. Define $W^{a,t}_2$ to be the event that the empirical mean of the first $t$ pulls of arm $a$ is at most $\sqrt{\frac{2\log(T)}{t}}$ away from $\mu_a$. Define $W_2$ to be the intersection of these events (i.e. $\bigcap_{a,t} W^{a,t}_2$).
By Chernoff bound,
\[
\Pr[W^{a,t}_2] \geq 1 - 2\exp(-4\log(T)) \geq 1-2/T^4.
\]
By union bound, 
\[
\Pr[W_2] \geq 1 - KT \cdot \frac{2}{T^4} \geq 1 - \frac{2}{T}.
\]
\end{itemize}

By union bound, we know $\Pr[W_1 \cap W_2] \geq 1 - 3/T$. Now we consider the case when $W_1 \cap W_2$ happens.

For any arm $a$ and agent $t$ in the second level, by $W_1$ and $W_2$, we have
\[
|\bar{\mu}^t_a - \mu_a| \leq \sqrt{\frac{2\log(T)}{q_aT_1 /2}}.
\]
By $W_1$ and Assumption \ref{ass:embehave}, we have
\[
|\bar{\mu}^t_a - \hat{\mu}^t_a| \leq \frac{c_m}{\sqrt{q_aT_1/2}}.
\]
Therefore,
\[
|\bar{\mu}^t_a - \mu_a|\leq \sqrt{\frac{2\log(T)}{q_aT_1 /2}}+\frac{c_m}{\sqrt{q_aT_1/2}} \leq 3 \sqrt{\frac{\log(T)}{\GdP T_1 }}
\]
So the second-level agents will pick an arm $a$ which has $\mu_a$ at most $6 \sqrt{\frac{\log(T)}{\GdP T_1 }}$ away from $\mu_1$. To sum up, the total expected regret is at most 
\[
T_1 \cdot \GdT + T \cdot (1-\Pr[W_1 \cap W_2]) + T \cdot  6 \sqrt{\frac{\log(T)}{\GdP T_1 }}.
\]
By setting $T_1 = T^{2/3}\log(T)^{1/3}$, we get expected regret $O(T^{2/3}\log(T)^{1/3})$.

Notice that if we know the gap between the means of the best arm and second best arm is larger than $\Delta$, we can set 
\[
T_1 = \max\left( \frac{4\GdT^2}{\GdP^2}\log(T), \frac{36 }{\Delta^2 \cdot p_G} \log(T) \right).
\]
In this case, since $\Delta \geq 6 \sqrt{\frac{\log(T)}{\GdP T_1 }}$, we know agents in the second level will all pull arm 1. Therefore, the total expected regret is at most
\[
T_1 \cdot \GdT + T \cdot (1- \Pr[W_1 \cap W_2]) = O(\Delta^{-2} \log(T)).
\]
\end{proof}

### 2018-10-02 llevel.tex
Update 2-level proof

Now we consider the case when $l > 2$ and assume the claim is true for smaller $l$'s. For each arm $a$, $W_3$ implies that there exists $u_a$ such that $W^{l-1,u_a,a,high}_3$ and $W^{l-1,u_a,3-a,low}_3$ happen. Recall $n^{l-1,u_a,a}$ is the number of arm $a$ pulls in groups $G_{l-1,u_a,1},...,G_{l-1,u_a,S}$. The induction hypothesis implies that $n^{l-1,u_a,a} \geq T_{l-1}$. $W^{l-1,u_a,a,high}_3$ together with $n^{l-1,u_a,a} \geq T_{l-1}$ implies that the empirical mean of arm $a$ pulls in group $G_{l-1,u_a,1},...,G_{l-1,u_a,S}$ is at least $\mu_a + 1/\sqrt{n^{l-1,u_a,a}}$. For any agent $t$ in groups $G_{l,1,u_a},...,G_{l,S,u_a}$ and $\Gamma_{l,1,u_a},...,\Gamma_{l,S,u_a}$, it observes history of groups $G_{l-1,u_a,1},...,G_{l-1,u_a,S}$ and all groups in levels below level $l-1$. Notice that the groups in the first $l-2$ levels have $(T_1 \GdT + T_2 + \cdots +T_{l-2})S^3 \leq T_{l-1}/(12\log(T)) \leq n^{l-1,u_a,a}/(12\log(T))$ agents. By $W_2$, we have
\begin{align*}
\bar{\mu}_a ^t &\geq \mu_a + \left(n^{l-1,u_a,a}  \cdot \frac{1}{\sqrt{n^{l-1,u_a,a} }}- (T_1 \GdT + T_2 + \cdots +T_{l-2})S^2\cdot \sqrt{\frac{3\log(T)}{ (T_1 \GdT + T_2 + \cdots +T_{l-2})S^2}} \right) \\
&~~~\cdot \frac{1}{n^{l-1,u_a,a}+ (T_1 \GdT + T_2 + \cdots +T_{l-2})S^2} \\
&> \mu_a + \frac{1}{4\sqrt{n^{l-1,u_a,a}  }}.
\end{align*}
The third last inequality holds when $T$ larger than some constant.
Similarly, we also have
\[
\bar{\mu}_{3-a}^t < \mu_{3-a}   -\frac{1}{4\sqrt{n^{l-1,u_a,3-a}  }}.
\]
Then we have
\begin{align*}
\bar{\mu}^t_a - \bar{\mu}^t_{3-a} &> \mu_a - \mu_{3-a}+ \frac{1}{4\sqrt{n^{l-1,u_a,a}  }} +\frac{1}{4\sqrt{n^{l-1,u_a,3-a}  }}\\
&\geq -\varepsilon_{l-1}+ \frac{1}{4\sqrt{n^{l-1,u_a,a}  }} +\frac{1}{4\sqrt{n^{l-1,u_a,3-a}  }}\\
&\geq \frac{1}{8\sqrt{n^{l-1,u_a,a}  }} +\frac{1}{8\sqrt{n^{l-1,u_a,3-a}  }}.
\end{align*}
The last inequality holds because $n^{l-1,u_a,a}$ and $n^{l-1,u_a,3-a}$ are at most $T_{l-1} S$. By Assumption \ref{ass:embehave}, we have
\begin{align*}
\hat{\mu}_a^t - \hat{\mu}_{3-a}^t &> \bar{\mu}^t_a - \bar{\mu}^t_{3-a} -  \frac{c_m}{\sqrt{n^{l-1,u_a,a} }} - \frac{c_m}{\sqrt{n^{l-1,u_a,3-a}}}\\
&> \frac{1}{8\sqrt{n^{l-1,u_a,a}  }} +\frac{1}{8\sqrt{n^{l-1,u_a,3-a}  }} -  \frac{c_m}{\sqrt{n^{l-1,u_a,a} }} - \frac{c_m}{\sqrt{n^{l-1,u_a,3-a}}}\\
&>0.
\end{align*}
The last inequality holds since $c_m$ is a small enough constant defined in Assumption \ref{ass:embehave}.
Therefore agents in groups $G_{l,1,u_a},...,G_{l,S,u_a}$ and $\Gamma_{l,1,u_a},...,\Gamma_{l,S,u_a}$ all pull arm $a$.
\end{proof}

>>>>

Now we consider the case when $l > 2$ and assume the claim is true for smaller $l$'s. For each arm $a$, $W_3$ implies that there exists $u_a$ such that $W^{l-1,u_a,a,high}_3$ and $W^{l-1,u_a,3-a,low}_3$ happen. Recall $n^{l-1,u_a,a}$ is the number of arm $a$ pulls in groups $G_{l-1,u_a,1},...,G_{l-1,u_a,S}$. The induction hypothesis implies that $n^{l-1,u_a,a} \geq T_{l-1}$. $W^{l-1,u_a,a,high}_3$ together with $n^{l-1,u_a,a} \geq T_{l-1}$ implies that the empirical mean of arm $a$ pulls in group $G_{l-1,u_a,1},...,G_{l-1,u_a,S}$ is at least $\mu_a + 1/\sqrt{n^{l-1,u_a,a}}$. For any agent $t$ in groups $G_{l,1,u_a},...,G_{l,S,u_a}$ and $\Gamma_{l,1,u_a},...,\Gamma_{l,S,u_a}$, it observes history of groups $G_{l-1,u_a,1},...,G_{l-1,u_a,S}$ and all groups in levels below level $l-1$. Notice that the groups in the first $l-2$ levels have $(T_1 \GdT + T_2 + \cdots +T_{l-2})S^2 \leq T_{l-1}/(12\log(T)) \leq n^{l-1,u_a,a}/(12\log(T))$ agents. By $W_2$, we have
\begin{align*}
\bar{\mu}_a ^t &\geq \mu_a + \left(n^{l-1,u_a,a}  \cdot \frac{1}{\sqrt{n^{l-1,u_a,a} }}- (T_1 \GdT + T_2 + \cdots +T_{l-2})S^2\cdot \sqrt{\frac{3\log(T)}{ (T_1 \GdT + T_2 + \cdots +T_{l-2})S^2}} \right) \\
&~~~\cdot \frac{1}{n^{l-1,u_a,a}+ (T_1 \GdT + T_2 + \cdots +T_{l-2})S^2} \\
&> \mu_a + \frac{1}{4\sqrt{n^{l-1,u_a,a}  }}.
\end{align*}
The third last inequality holds when $T$ larger than some constant.
Similarly, we also have
\[
\bar{\mu}_{3-a}^t < \mu_{3-a}   -\frac{1}{4\sqrt{n^{l-1,u_a,3-a}  }}.
\]
Then we have
\begin{align*}
\bar{\mu}^t_a - \bar{\mu}^t_{3-a} &> \mu_a - \mu_{3-a}+ \frac{1}{4\sqrt{n^{l-1,u_a,a}  }} +\frac{1}{4\sqrt{n^{l-1,u_a,3-a}  }}\\
&\geq -\varepsilon_{l-1}+ \frac{1}{4\sqrt{n^{l-1,u_a,a}  }} +\frac{1}{4\sqrt{n^{l-1,u_a,3-a}  }}\\
&\geq \frac{1}{8\sqrt{n^{l-1,u_a,a}  }} +\frac{1}{8\sqrt{n^{l-1,u_a,3-a}  }}.
\end{align*}
The last inequality holds because $n^{l-1,u_a,a}$ and $n^{l-1,u_a,3-a}$ are at most $T_{l-1} S$. By Assumption \ref{ass:embehave}, we have
\begin{align*}
\hat{\mu}_a^t - \hat{\mu}_{3-a}^t &> \bar{\mu}^t_a - \bar{\mu}^t_{3-a} -  \frac{c_m}{\sqrt{n^{l-1,u_a,a} }} - \frac{c_m}{\sqrt{n^{l-1,u_a,3-a}}}\\
&> \frac{1}{8\sqrt{n^{l-1,u_a,a}  }} +\frac{1}{8\sqrt{n^{l-1,u_a,3-a}  }} -  \frac{c_m}{\sqrt{n^{l-1,u_a,a} }} - \frac{c_m}{\sqrt{n^{l-1,u_a,3-a}}}\\
&>0.
\end{align*}
The last inequality holds since $c_m$ is a small enough constant defined in Assumption \ref{ass:embehave}.
Therefore agents in groups $G_{l,1,u_a},...,G_{l,S,u_a}$ and $\Gamma_{l,1,u_a},...,\Gamma_{l,S,u_a}$ all pull arm $a$.
\end{proof}

### 2018-10-02 model.tex
Update 2-level proof

\emph{Bayesian Exploration} is a game between a principal and $T$ agents who arrive one by one. Each round $t$ proceeds as follows: a new agent $t$ arrives, receives a message $m_t$ from the principal, chooses an action $a_t$ from a fixed action space $\A = \{1,...,K\}$, and collects a reward $r_t\in \{0,1\}$ that is immediately observed by the principal.

Each action $a \in \A$ has a state $\mu_a$ drawn from $\D_a$ before the game starts. For agent in round $t$ who takes action $a_t$, the reward $r_t$ is distributed as $\Ber(\mu_{a_t})$.

@@

\xhdr{Restrictions on the messages.} We classify recommendation policies according to what messages are allowed. 
\begin{itemize}
\item Full History: 
\item Unbiased Sub-history: 
\end{itemize}

@@

\xhdr{Agents behavior.} We assume agent $t$ estimates the mean of each arm as $\hat{\mu}_a^t$'s and pick the arm with the highest $\hat{\mu}_a^t$'s. We also define $\bar{\mu}_a^t$ as the empirical average of arm $a$ given the history. We assume each agent's estimate $\hat{\mu}_a^t$ is close to the empirical average $\bar{\mu}_a^t$, more specifically as the following assumption:
\begin{assumption}
\label{ass:embehave}
Let constant $c_m = 1/16$. There exists a constant $c_T$ such that the following is true. For each agent $t$ and arm $a$, let $t_a$ be the number of times arm $a$ has been pulled in the history observed by agent $t$. 
\begin{itemize}
\item \textbf{Close to the empirical average:} If $t_a \geq c_T$, 
\[
|\hat{\mu}^t_a - \bar{\mu}^t_a | < \frac{c_m}{\sqrt{t_a}}.
\]
\item \textbf{Initial value:} If $t_a = 0$, $\hat{\mu}^t_a$ is at least $1/3$. 
\end{itemize}
\end{assumption}

We allow different agents to have different behavior models as long as they satisfy Assumption \ref{ass:embehave}. More specifically, we assume each agent's behavior model is sampled from an unknown distribution independently.

>>>>

\emph{Incentivizing Exploration} is a game between a principal and $T$ agents who arrive one by one. Each round $t$ proceeds as follows: a new agent $t$ arrives, receives a message $m_t$ from the principal, chooses an arm $a_t$ from a fixed arm space $\A = \{1,...,K\}$, and collects a reward $r_t\in \{0,1\}$ that is immediately observed by the principal. 
We mostly focus on the case when $K=2$. We also discuss how to generalize our results to the case when $K$ is a constant greater than 2.

Each arm $a \in \A$ has a state $\mu_a$ drawn from $\D_a$ before the game starts. For the agent in round $t$ who picks arm $a_t$, the reward $r_t$ is distributed as $\Ber(\mu_{a_t})$.

@@

\xhdr{Restrictions on the messages.} We classify recommendation policies according to what messages are allowed. 
\begin{itemize}
\item \textbf{Full history:} We show each agent the history of all previous agents' pulls. We call this policy \ALGG. 
\item \textbf{Unbiased sub-history:} We show each agent only a subset of history. The subsets are pre-determined before any arms are pulled.  
\end{itemize}

@@

\xhdr{Agents' behavior.} We assume agent $t$ estimates the mean of each arm as $\hat{\mu}_a^t$'s and picks the arm with the highest $\hat{\mu}_a^t$. We define $\bar{\mu}_a^t$ as the empirical average of arm $a$ given the history. We assume each agent's estimate $\hat{\mu}_a^t$ is close to the empirical average $\bar{\mu}_a^t$, more specifically as the following assumption:
\begin{assumption}
\label{ass:embehave}
Set constant $c_m = 1/16$. There exists a constant $c_T$ such that the following is true. For each agent $t$ and arm $a$, let $t_a$ be the number of times arm $a$ has been pulled in the history observed by agent $t$. 
\begin{itemize}
\item \textbf{Close to the empirical average:} If $t_a \geq c_T$, 
\[
|\hat{\mu}^t_a - \bar{\mu}^t_a | < \frac{c_m}{\sqrt{t_a}}.
\]
\item \textbf{Initial value:} If $t_a = 0$, $\hat{\mu}^t_a$ is at least $1/3$. 
\end{itemize}
\end{assumption}

We allow different agents to have different behavior models as long as they satisfy Assumption \ref{ass:embehave}. More specifically, we assume each agent's behavior model is sampled from an unknown distribution independently. 
\begin{remark}
how this relates to Bayesian exploration with conjugate priors
\end{remark}
